---
title: "SDS 293: Modeling for Machine Learning"
author: "Albert Y. Kim"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: kable
    includes:
      in_header: "favicon.html"
---

<style>
h1{font-weight: 400;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE, 
  message=FALSE, 
  warning=FALSE, 
  eval=FALSE, 
  fig.width=16/2, 
  fig.height=9/2
)
library(tidyverse)
library(broom)
library(knitr)
library(modelr)
library(lubridate)
library(forcats)
library(nycflights13)
library(patchwork)
library(okcupiddata)
# devtools::install_github("hadley/emo")
library(emo)

# Set seed value of random number generator to get "replicable" random numbers.
set.seed(76)
```

```{r, eval=FALSE, echo=FALSE}
# Run this separately to have slide output:
rmarkdown::render("index.Rmd", output_format = c("ioslides_presentation"), output_file = "slides.html")
```

<style>
h1{font-weight: 400;}
</style>



***



# Schedule 

<iframe src="https://docs.google.com/spreadsheets/d/e/2PACX-1vToPV6gfDlIq5ni1qezMcPy3ZdyIN1MtSSKZ3GABeBvf9LYu3_1XE7DOEQh4Dg02bKG5YF0XpSFO_-B/pubhtml?gid=2021662517&amp;single=true&amp;widget=true&amp;headers=false" width="100%" height="775"></iframe>

<!--
{target="_blank"}
-->



***



# Lec 03: Fri 1/31

## Announcements

* GitHub organization for this class:
    + Access it by clicking the "octocat" icon on the top left of this page
    + Ensure you are a member (I sent an email invite yesterday)
* GitHub profile. Think of it as an extension of your resume. I highly encourage, but do not require, you to:
    + Post your full name
    + Post your affiliation
    + Post a public-facing profile picture (it doesn't have to be an image of you, it can be any picture).
    + Please don't change your GitHub ID's mid-semester!
* FYI: GitHub is not without [controversy](https://techcrunch.com/2019/11/13/github-faces-more-resignations-in-light-of-ice-contract/){target="_blank"}

## Topics

GitHub has many definitions that are unforunately not straight forward. Using [`fivethirtyeight`](https://github.com/rudeboybert/fivethirtyeight){target="_blank"} R package as an example.

* Git vs GitHub
* Local vs remote
* Repos & `README.md` files as cover pages
* Cloning a repo locally (which in our case correspond to RStudio Projects). Ex: Big green "Clone or download" button on top right.
* Commit/push and pull. Commits as units of change. Ex: Click "commits" tab on top left (currently at 523 commits)
* Branches: `master` branch is what you see. Ex: Click "Branch" button on top left 
* Making contributions to `master`
    + When on the inside: Create a new branch, make edits in branch, then make pull request
    + When on the outside: Fork the repo (similar to creating a branch), make edits in your forked copye of repo, then make pull request
    + Ex: Click on Insights tab on top right -> Network -> Scroll to earlier dates. 
* Learning git has a steep learning curve. This is normal:
    + [xkcd](https://xkcd.com/1597/){target="_blank"}
    + [Oh shit, git!](https://gumroad.com/l/oh-shit-git){target="_blank"}
    + In this class: I'll (gently) force you to practice merge conflicts




***



# Lec 02: Wed 1/29

## Announcements

* On Sun 1/26 I sent an email including links to join Slack and a Intro Survey Google Form. If you did not receive this email, please come see me after class.
* Syllabus [posted](syllabus.html)


## Topics

* In [MassMutual Google Doc](https://docs.google.com/document/d/1kBl8ivTeNyVfbV0gvXjBpwnlRWXvZkKuSe0SMTjFB38/edit#heading=h.4llig2jamgry){target="_blank"}: Tuesday AM.



***



# Lec 01: Mon 1/27

## Announcements


## Topics

* Went over "What is Machine Learning?" [slides](static/what_is_ML.pdf){target="_blank"}.
