---
title: "SDS 293: Modeling for Machine Learning"
author: "Albert Y. Kim"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    df_print: kable
    includes:
      in_header: "favicon.html"
---

<style>
h1{font-weight: 400;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE, 
  message=FALSE, 
  warning=FALSE, 
  eval=FALSE, 
  fig.width=16/2, 
  fig.height=9/2
)
library(tidyverse)
library(broom)
library(knitr)
library(modelr)
library(lubridate)
library(forcats)
library(nycflights13)
library(patchwork)
library(okcupiddata)
# devtools::install_github("hadley/emo")
library(emo)

# Set seed value of random number generator to get "replicable" random numbers.
set.seed(76)
```

```{r, eval=FALSE, echo=FALSE}
# Run this separately to have slide output:
rmarkdown::render("index.Rmd", output_format = c("ioslides_presentation"), output_file = "slides.html")
```

<style>
h1{font-weight: 400;}
</style>



***



# Schedule 

<iframe src="https://docs.google.com/spreadsheets/d/e/2PACX-1vToPV6gfDlIq5ni1qezMcPy3ZdyIN1MtSSKZ3GABeBvf9LYu3_1XE7DOEQh4Dg02bKG5YF0XpSFO_-B/pubhtml?gid=2021662517&amp;single=true&amp;widget=true&amp;headers=false" width="100%" height="775"></iframe>

<!--
{target="_blank"}
-->



***



# Lec 03: Fri 1/31

## Announcements

* 


## Topics

* GitHub profile. Think of it as an extension of your resume. I highly encourage (but do not require)
    + You post your full name
    + A profile picture that you are comfortable being public facing. It doesn't have to be an image of you, it can be an image.
    + Affiliation
* Terminology:
    + Git vs GitHub. Other online interfaces to Git exist, like BitBucket and GitLab. GitHub is not without [controversy](https://techcrunch.com/2019/11/13/github-faces-more-resignations-in-light-of-ice-contract/){target="_blank"}. 
    + Cloning a repo. In our case, they correspond to RStudio Project folders
    + local vs remote
    + commit/push and pull. Commits as units of change.
    + master branch
    + When on the inside: branch + pull request
    + When on the outside: fork + pull request
    + Oh shit, git! Ex: merge conflicts.
* GitHub Classroom:
    + Layer around standard GitHub organization
    + Based on [GitHub organization](https://github.com/293-2020-01-Fall){target="_blank"}. Accept membership please!



***



# Lec 02: Wed 1/29

## Announcements

* On Sun 1/26 I sent an email including links to join Slack and a Intro Survey Google Form. If you did not receive this email, please come see me after class.
* Syllabus [posted](syllabus.html)


## Topics

* In [MassMutual Google Doc](https://docs.google.com/document/d/1kBl8ivTeNyVfbV0gvXjBpwnlRWXvZkKuSe0SMTjFB38/edit#heading=h.4llig2jamgry){target="_blank"}: Tuesday AM.



***



# Lec 01: Mon 1/27

## Announcements


## Topics

* Went over "What is Machine Learning?" [slides](static/what_is_ML.pdf){target="_blank"}.
